---
title: "Intubation - An extact of the SIMULRUN project data"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

###The dataset called intubation is a sample of SIMULRUN's data. SIMULRUN is a research project funded by European funds (FEDER) and focusses on learning processes in simulation in healthcare. This set of data corresponds to an intubation procedure of a new-born resuscitation during a high fidelity simulation.

## Contruction of the ViSibook for procedure
```{r, fig.show='hold',fig.width=7}
vars <- c("time_in_intub","time_insert_probe","time_out_intub","delay_intub_prob") 
label <- c( "Blade in the mouth", "Tube insert in the mouth",
           "Blade out of the mouth","Blade in - Tube inserted")
typeA <- c( "p","p","p","l")
showorder <- c( 1,3,4,2)
deb   <- c( NA, NA,NA,"time_in_intub")
fin <- c( NA,NA,NA,"time_insert_probe")
bookdataframe <- data.frame(vars,label,typeA,showorder,deb,fin)
library(ViSiElse)
bookintubation <- ConvertoViSibook(bookdataframe)
plot(bookintubation)
```

## Graphic of the procedure  
```{r, fig.show='hold',fig.width=7,fig.height=7}
x <- buildViSiGrid( X = intubation, 
                              book = bookintubation ,
                              pixel = 2 )
plot(x, main = "Intubation, t_0 : Blade in the mouth")
```

## Graphic of the procedure strating when the blade is in the mouth
```{r, fig.show='hold',fig.width=7, fig.height=7}
bookintubation <- changeShoworder( bookintubation , c(2,3,4))
Vintubation <- buildViSiGrid( X = intubation, 
                              book = bookintubation ,
                              t_0 = "time_in_intub",pixel = 1 )
plot(Vintubation, main = "Intubation, t_0 : Blade in the mouth")
```



------
### Any question?
If you have any request or comment >> https://github.com/CEPOI/ViSiElse


...

